---
title: "get_precip"
output: html_document
author: Corinne Bowers
date: 2022-09-27
theme: spacelab   #https://www.datadreaming.org/post/r-markdown-theme-gallery/
highlight: tango  #https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/
---

This script creates a raster stack of precipitation, measured as total mm of accumulation over the specified time period (either 3 hours or 24 hours).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/scratch/users/cbowers/sequences/')

```

# setup 

## functions & packages 

```{r}
source('_data/setup.R')
cores <- parallel::detectCores()-2

## define function to convert from array to raster
array_to_raster <- function(x) {
  x %>% 
    aperm(c(2,1,3)) %>% 
    brick(
      xmn = min(lon_merra)-xres/2, xmx = max(lon_merra)+xres/2,
      ymn = min(lat_merra)-yres/2, ymx = max(lat_merra)+yres/2,
      crs = projection(wus)) %>%
    raster::flip('y')
}

```

## metadata

```{r}
## define MERRA timeseries
ts_merra <- 
  seq(ymd_hms('1980-1-1 00:00:00'), ymd_hms('2021-12-31 21:00:00'), by = '3 hours')
dates_merra <- ts_merra %>% as.Date %>% unique

## define MERRA lat/lon bounding box
lon_merra <- seq(-140, -100, xres) + xres/2
lat_merra <- seq(20, 60, yres) + yres/2

## define empty raster
empty <- grid_ca %>% setValues(NA)

```

# download MERRA-2 precipitation files

The process for downloading the MERRA-2 hourly precipitation files is as follows: 

1. Get a list of wget links from https://disc.gsfc.nasa.gov/ and save the list as `url.txt`. You need an Earthdata account to do this. 
  + The correct file should be M2T1NXFLX.5.12.4. Make sure to download MERRA-2, not MERRA, and check that the spatial resolution is 0.5x0.625 and the temporal resolution is one hour. 
  + Select the correct variables (we used PRECTOT and PRECTOTCORR) and the correct bounding box (we used [-140, 20, -100, 60] to capture the western US). 
  + Please note that the link list will expire within a few days of generation.
2. Run `wget_loop.sh` to iteratively download files, one per day. Check in occasionally and restart the process when it stalls.
3. Run the code chunk below to identify failed file downloads. 
4. Iterate steps 2-3 until a sufficient number of files have been successfully downloaded.

```{r}
## check for failed file dates
files <- list.files('/scratch/users/cbowers/MERRA/files') %>% .[grepl('MERRA2_', .)]
filedates <- files %>% str_split('\\.') %>% lapply(function(x) x[[3]]) %>% unlist %>% ymd
dates <- dates_merra
match(dates, filedates) %>% is.na %>% which %>% dates[.]

```

# create precipitation raster

## 3-hour data

```{r}
## load all 3-hour precipitation values
start <- Sys.time()
cl <- parallel::makeCluster(cores)
registerDoSNOW(cl)
pb <- txtProgressBar(min = 0, max = length(dates), style = 3)
PRCP <- 
  foreach (
    i = 1:length(dates),
    .export = 'array_to_raster',
    .options.snow = opts,
    .packages = c('ncdf4', 'raster', 'sf', 'dplyr', 'lubridate')) %dopar% {
      if (dates[i] %in% filedates) { ## add precipitation info to daily list
        nc <- nc_open(paste0('/scratch/users/cbowers/MERRA/files/', files[which(dates[i] == filedates)]))
        prcp <- ncvar_get(nc, 'PRECTOT') / 1e3 * 60^2 * 1e3  #mm/hr
        nc_close(nc)
        array_to_raster(prcp) %>% 
          crop(grid_ca) %>% 
          stackApply(indices = rep(1:8, each = 3), fun = sum)  #mm total in 3 hours 
      } else stack(mget(rep('empty',8))) ## leave blank if data is missing
    }
stopCluster(cl)

## convert to raster stack
prcp <- PRCP %>% do.call('stack', .)

## save out
save(prcp, file = '_data/MERRA/precip/precip_3hr_0922.Rdata')
Sys.time() - start

```

## 24-hour data 

```{r}
## load all daily precipitation values
start <- Sys.time()
cl <- parallel::makeCluster(cores)
registerDoSNOW(cl)
pb <- txtProgressBar(min = 0, max = length(dates), style = 3)
PRCP <- 
  foreach (
    i = 1:length(dates),
    .export = 'array_to_raster',
    .options.snow = opts,
    .packages = c('ncdf4', 'raster', 'sf', 'dplyr', 'lubridate')) %dopar% {
      if (dates[i] %in% filedates) { ## add precipitation info to daily list
        nc <- nc_open(paste0('../MERRA/files/', files[which(dates[i] == filedates)]))
        prcp <- ncvar_get(nc, 'PRECTOT') / 1e3 * 60^2 * 1e3  #mm/hr
        nc_close(nc)
        array_to_raster(prcp) %>% 
          crop(grid_ca) %>% 
          calc(sum)  #mm daily total 
      } else empty  ## leave blank if data is missing
    }
stopCluster(cl)

## convert to raster stack
prcp <- PRCP %>% do.call('stack', .)

## save out
save(prcp, file = '_data/MERRA/precip/precip_24hr_0928.Rdata')
Sys.time() - start

```



