---
title: "create_df"
output: html_document
date: "2022-09-27"
author: Corinne Bowers
theme: spacelab   #https://www.datadreaming.org/post/r-markdown-theme-gallery/
highlight: tango  #https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/
---

This script creates three dataframes for one of the preprocessed ARDTs specified by the user. The ARDT options are:

* Brands et al. (2017), v3        # https://doi.org/10.1007/s00382-016-3095-6
* CASCADE - IVT                   # N/A
* CONNECT500                      # https://doi.org/10.1175/JHM-D-14-0101.1
* Gershunov et al. (2017)         # https://doi.org/10.1002/2017GL074175
* Goldenson et al. (2018), v1     # https://doi.org/10.1175/JCLI-D-18-0268.1
* Guan & Waliser (2015)           # https://doi.org/10.1002/2015JD024257
* LBNL ML-TDA                     # https://doi.org/10.5194/gmd-12-613-2019
* Lora et al. (2017), N. Pacific  # https://doi.org/10.1002/2016GL071541
* Mattingly et al. (2018), v2     # https://doi.org/10.1029/2018JD028714
* Mundhenk et al. (2016), v2      # https://doi.org/10.1175/JCLI-D-15-0655.1
* Payne & Magnusdottir (2016)     # https://doi.org/10.1002/2016JD025549
* Rutz et al. (2014)              # https://doi.org/10.1175/MWR-D-13-00168.1
* SCAFET, v1                      # N/A
* Walton                          # N/A

The dataframes are: a list of three-hour datasets, a list of twenty-four-hour datasets, and list of event catalogs, separated by MERRA-2 grid cell. 


```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = 'D:/02-sequences/')
# knitr::opts_knit$set(root.dir = '/scratch/users/cbowers/sequences/')
# knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(results = 'hold', fig.show = 'hold', fig.align = 'center')
# rm(list=ls())

```

# setup 

```{r}
## load packages & functions
source('_data/setup_impacts.R')
source('_scripts/create_df_functions.R')
# cores <- parallel::detectCores()-2

mode <- function(x) ifelse(all(is.na(x)), NA, names(which.max(table(x)))[1])

## turn progress bars on/off
progress <- TRUE
if (progress) pb <- txtProgressBar(min = 0, max = ncell(grid_ca), style = 3)

## set number of bootstrapped samples
boot <- 1000

## set random seed for reproducibility
set.seed(2023)

```

## choose ARDT

```{r}
ardtnames <- c(
  'brands','cascade','connect','gershunov','goldenson','guanwaliser','lbnl',
  'lora','mattingly','mundhenk','paynemagnusdottir','rutz','scafet','walton')
ardtname <- 'lbnl'

```

## load data

### metadata

```{r}
## get hourly/daily timeseries
ts_merra <- 
  seq(ymd_hms('1980-1-1 00:00:00'), ymd_hms('2021-12-31 21:00:00'), by = '3 hours')
dates_merra <- ts_merra %>% as.Date %>% unique

ts_artmip <- 
  seq(ymd_hms('1980-1-1 00:00:00'), ymd_hms('2016-12-31 21:00:00'), by = '3 hours')

```

### existing IVT dataframes

```{r}
load('_scripts/_checkpoints/df_3hr_1209.Rdata')

```

### ARs

```{r}
load(paste0('_data/ARTMIP/Tier 1/ardt_',ardtname,'.Rdata'))

```

### impacts

```{r}
load('_data/impacts/ncei_0222.Rdata')         # NCEI storm events database
load('_data/impacts/NFIP_0210.Rdata')         # NFIP claims & policies

```

## define functions

```{r}
add_inter <- function(df) {
  #' Creates the colum inter.duration measuring between-event interval, in days.
  #' @param df dataframe with columns ts (datetime), ar (logical), and ivt (double)
  #' @return dataframe with new column inter.duration (double)
  temp <- df %>% mutate(inter.count = add_counter(inter))
  temp <- create_catalog(temp, 'inter', cat = FALSE, interval = 3/24) %>% 
    select(count, duration) %>% 
    setNames(paste('inter', names(.), sep = '.')) %>% 
    left_join(temp, ., by = 'inter.count') %>% 
    mutate(
      inter.duration = setNA(inter.duration,0),
      prev.inter = c(NA, inter.duration[-nrow(.)]),
      next.inter = c(inter.duration[-1], NA)) %>% 
    select(c(names(temp), ends_with('inter'))) %>% select(-inter, -inter.count)
  return(temp)
}

add_category <- function(df) {
  #' Adds the column ar.cat to a 3-hour timeseries dataframe.
  #' @param df dataframe with columns count (double), ivt (double), and duration (double)
  #' @return dataframe with new column ar.cat
  temp <- df %>% 
    filter(!is.na(count)) %>% 
    group_by(count) %>% 
    summarize(duration = max(duration), maxivt = max(ivt)) %>% 
    filter(duration > 6) %>% 
    mutate(ar.cat = map2_dbl(.x = maxivt, .y = duration, .f = ~assign_AR_cat(.x,.y)))
  temp <- df %>% left_join(temp %>% select(count, ar.cat), by = 'count')
}

mode <- function(x) ifelse(all(is.na(x)), NA, names(which.max(table(x)))[1])

```

# create dataframes

## create 3-hr dataframes 

```{r}
## convert from list by year to list by location
start <- Sys.time()
cl <- makeCluster(cores)
registerDoSNOW(cl)
artmip_3hr <-
  foreach(
    i = 1:ncell(grid_ca),
    .options.snow = if (progress) opts,
    .packages = c('raster', 'foreach', 'tidyverse', 'lubridate'),
    .export = c('assign_AR_cat', 'add_counter', 'add_category')) %dopar% {
      if (i %in% index_ca) {
        index <- rowColFromCell(grid_ca, i)
        data <-
          data.frame(
            ts = ts_artmip, #have to do ts_artmip[-1] for SCAFET
            ar = ardt[1:37] %>%
              lapply(function(x) c(x[index[1],index[2],])) %>%
              reduce(c) %>% round %>% as.logical) %>% 
          mutate(count = add_counter(ar)) %>% 
          left_join(df_3hr[[i]] %>% select(ts,ivt,seq,seq.count), by = 'ts')
        catalog <-
          create_catalog(data, name = 'ar', interval = 3) %>%
          filter(duration > 6) %>%
          mutate(start = paste(start), end = paste(end))
        data %>%
          rename(ar.count = count) %>% 
          left_join(
            catalog %>% setNames(paste('ar', names(.), sep = '.')),
            by = 'ar.count') %>%
          mutate(
            ar = ifelse(!is.na(ar.duration), TRUE, FALSE),
            ar.count = ifelse(!is.na(ar.duration), ar.count, NA)) %>% 
        # data %>% 
        #   calculate_duration(.) %>% add_category(.) %>% 
        #   rename(ar.count = count) %>% 
          mutate(inter = case_when(ar.cat==0 ~ TRUE, TRUE ~ !ar)) %>%
          add_inter(.)
      } else NULL
    }
stopCluster(cl)
Sys.time() - start

```

## create 24-hr dataframes

```{r}
start <- Sys.time()
cl <- makeCluster(cores)
registerDoSNOW(cl)
artmip_24hr <-
  foreach(
    i = 1:ncell(grid_ca),
    .options.snow = if (progress) opts,
    .packages = c('raster', 'foreach', 'tidyverse', 'lubridate'),
    .export = c('mode', 'setNA')) %dopar% {
      # if (progress) setTxtProgressBar(pb,i)
      if (i %in% index_ca) {
        artmip_3hr[[i]] %>% 
          group_by(date = as.Date(ts)) %>% 
          summarize(
            ar = mean(ar) >= 0.25,
            ar.count = case_when(ar ~ mode(ar.count)),
            seq = mean(seq) >= 0.25) %>% 
            left_join(
              ncei_grid %>% filter(cell == i) %>% select(-cell),
              by = 'date') %>%
            mutate(
              ncei_damage = case_when(year(date) >= 1996 ~ setNA(ncei_damage,0)),
              ncei_event = case_when(year(date) >= 1996 ~ setNA(ncei_event, FALSE))) %>% 
            left_join(
              claims_grid %>% filter(cell == i) %>% select(-cell),
              by = 'date') %>%
            mutate(
              claims_num = setNA(claims_num,0),
              claims_value = setNA(claims_value,0)) %>% 
          select(-claims_coverage, -claims_dr)
      } else NULL
    }
stopCluster(cl)
Sys.time() - start

```

## create event catalog 

```{r}
start <- Sys.time()
# cl <- makeCluster(cores)
# registerDoSNOW(cl)
artmip_catalog <-
  foreach(
    i = 1:ncell(grid_ca),
    .options.snow = opts,
    .packages = c('raster', 'foreach', 'tidyverse', 'lubridate'),
    .export = c('toNumber')) %do% {
      if (progress) setTxtProgressBar(pb,i)
      if (i %in% index_ca) {
        artmip_3hr[[i]] %>% 
          ###
          mutate(
            ar = case_when(ar.cat==0 ~ FALSE, TRUE ~ ar),
            ar.cat = case_when(ar ~ ar.cat),
            ar.count = case_when(ar ~ ar.count)) %>% 
          ###
          filter(!is.na(ar.count)) %>% select(-ar) %>% 
          group_by(ar.count) %>% 
          summarize(
            across(c(starts_with('ar'), starts_with('seq')), ~.[1]),
            prev.inter = prev.inter[1],
            next.inter = next.inter[length(ar.count)]) %>% 
          mutate(
            prev.cat = c(NA, ar.cat[-nrow(.)]),
            next.cat = c(ar.cat[-1], NA),
            prev.inter = c(NA, prev.inter[-1]), 
            next.inter = c(next.inter[-nrow(.)], NA)) %>% 
          left_join(
            artmip_24hr[[i]] %>% 
              filter(!is.na(ar.count)) %>% 
              group_by(ar.count = toNumber(ar.count)) %>% 
              summarize(across(c(starts_with('ncei'), starts_with('claims')), Sum)),
            by = 'ar.count') %>% 
          mutate(id = i)
      } else NULL
    }
# stopCluster(cl)
Sys.time() - start

```

## save out

```{r}
## save out
save(
  artmip_3hr, artmip_24hr, artmip_catalog,
  file = paste0('_data/ARTMIP/Tier 1/df_',ardtname,'.Rdata'))

```

# calculate coefficients

## load data

```{r}
# load(paste0('_data/ARTMIP/Tier 1/df_',ardtname,'.Rdata'))

df_3hr <- artmip_3hr
df_24hr <- artmip_24hr
df_catalog <- artmip_catalog

```

## Create AR event catalog 

```{r}
## keep WY 1981-2021 (for category analysis)
catalog.full <- df_catalog %>% 
  reduce(rbind) %>% 
  mutate(wy = wateryear(ar.start)) %>%
  filter(wateryear(ar.start) %in% 1981:2021)
  
## keep WY 1997-2021 (for impact analysis)
catalog <- catalog.full %>%
  filter(wateryear(ar.start) %in% 1997:2021) %>%
  mutate(
    ncei_damage = setNA(ncei_damage,0), 
    claims_value = setNA(claims_value,0),
    ncei_nonzero = ncei_event > 0,  # %%%%%%%%%%%%%%%%% should this be different?
    nfip_nonzero = claims_num > 0) # %%%%%%%%%%%%%%%%% should this be different?

```

## Result #2: Sequences and likelihood of impact

```{r}
daily <- 
  map_dfr(
    .x = index_ca, 
    .f = ~df_24hr[[.x]] %>% mutate(id = .x)) %>% 
  mutate(
    impactday = claims_num>0 | ncei_event,
    lossday = claims_value>0 | ncei_damage>0) %>% 
  mutate(wy = wateryear(date)) %>% 
  filter(wy %in% 1997:2021)

f2 <- felm(
  impactday ~ ar*seq - seq | factor(id) + factor(wy), 
  data = daily, exactDOF = TRUE)
impact <- coef(summary(f2)) %>% data.frame

```

## Result #3: Sequences and magnitude of impact

```{r}
## NCEI
index.ncei.zero <- which(catalog$ncei_damage == 0)
index.ncei.nonzero <- which(catalog$ncei_damage > 0)
coef.under <-
  foreach (b = 1:boot) %do% {
    felm(
      log10(ncei_damage+1) ~ factor(ar.cat) + seq | factor(id) + factor(wy),
      data = catalog %>% 
        slice(c(index.ncei.nonzero, sample(
          index.ncei.zero, size = 2*length(index.ncei.nonzero), replace = FALSE))),
      exactDOF = TRUE) %>%
      summary %>% coef
  } %>% 
  lapply(function(x) {
    if (dim(x)[1] < 5) {
      x %>% data.frame %>% rownames_to_column %>%
        full_join(data.frame(rowname = c(paste0('factor(ar.cat)',2:5),'seqTRUE'))) %>%
        arrange(rowname) %>% 
        column_to_rownames %>% as.matrix
    } else x
  }) %>% abind(along = 3) %>% apply(c(1,2),Mean)
df.ncei <- coef.under %>% data.frame %>%
  rownames_to_column %>% 
  transmute(
    cat = c(gsub('factor\\(ar\\.cat\\)','',rowname[-length(rowname)]),NA),
    seq = c(rep(NA, nrow(.)-1),TRUE),
    est = 10^Estimate, 
    lower = 10^(Estimate-2*Std..Error), 
    upper = 10^(Estimate+2*Std..Error)) %>%
  rbind(data.frame(cat = 1, seq = NA, est = 1, lower = 1, upper = 1))

## NFIP
index.nfip.zero <- which(catalog$claims_value == 0)
index.nfip.nonzero <- which(catalog$claims_value > 0)
coef.under <- 
  foreach (b = 1:boot) %do% {
    felm(
      log10(claims_value+1) ~ factor(ar.cat) + seq | factor(id) + factor(wy),
      data = catalog %>% 
        slice(c(index.nfip.nonzero, sample(
          index.nfip.zero, size = 2*length(index.nfip.nonzero), replace = FALSE))),
      exactDOF = TRUE) %>% 
      summary %>% coef
  } %>% 
  lapply(function(x) {
    if (dim(x)[1] < 5) {
      x %>% data.frame %>% rownames_to_column %>%
        full_join(data.frame(rowname = c(paste0('factor(ar.cat)',2:5),'seqTRUE'))) %>%
        arrange(rowname) %>% 
        column_to_rownames %>% as.matrix
    } else x
  }) %>% abind(along = 3) %>% apply(c(1,2),Mean)
df.nfip <- coef.under %>% data.frame %>% 
  rownames_to_column %>% 
  transmute(
    cat = c(gsub('factor\\(ar\\.cat\\)','',rowname[-length(rowname)]),NA),
    seq = c(rep(NA, nrow(.)-1),TRUE),
    est = 10^Estimate, 
    lower = 10^(Estimate-2*Std..Error), 
    upper = 10^(Estimate+2*Std..Error)) %>% 
  rbind(data.frame(cat = 1, seq = NA, est = 1, lower = 1, upper = 1))

```

## Save coefficient estimates

```{r}
save(
  impact, df.ncei, df.nfip, 
  file = paste0('_data/ARTMIP/Tier 1/coef_',ardtname,'.Rdata'))

```

